---
title: "R² and Pseudo R²"
author: "Rodolfo Pelinson"
date: "2024-10-11"
output: github_document
---
We first will simulate a linear effe between two variables.

```{r}
x <- runif(50, min = 0, max = 100)
y <- rnorm(50, mean = 10 + 1.5*x, sd = 30)

plot(y ~ x)
```

Now lets compute the regular coefficient of determination
```{r}
model <- lm(y~x)
model0 <- lm(y~1)

regular_r2 <- summary(model)$r.squared
mcfadden_r2 <- c(1 - (logLik(model)/logLik(model0)))
LR_r2 <- (deviance(model0) - deviance(model))/ deviance(model0)
```


Now lets simulate some scenarios and see how things vary according to a large variation in r² values.

```{r}
sds <- runif(50, max = 100, min = 1)

regular_r2 <- rep(NA, 50)
mcfadden_r2 <- rep(NA, 50) 
LR_r2 <- rep(NA, 50)
  
for (i in 1:50){
  x <- runif(50, min = 0, max = 100)
  y <- rnorm(50, mean = 10 + 1.5*x, sd = sds[i])
  
  model <- lm(y~x)
  model0 <- lm(y~1)

  regular_r2[i] <- summary(model)$r.squared
  mcfadden_r2[i] <- c(1 - (logLik(model)/logLik(model0)))
  LR_r2[i] <- (deviance(model0) - deviance(model))/ deviance(model0)
}

plot(mcfadden_r2 ~ regular_r2)
plot(LR_r2 ~ regular_r2)
```

Likelyhood R² seems to make a better job in terms of linear scaling and similarity to absolute values. But it has a problem of having a maximum of less than 1 for logistic regression. the function r.squaredLR from the MuMIn package offers a correction.

